A Combinatorial Search of Parameterized Quantum Circuit Learning for Chemical Applications

### 2025-01-30 manuscript/main.tex
Thu Jan 30 04:52:31 PM EST 2025 manuscript images

Function fitting 5: 
Function fitting 16: all ran with 1000 iterations
BSE 5
BSE 16
DDCC

@@

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/sixteenqubit/16qubit_Linear_RUD_AL}
		\caption{}
		\label{fig:16qubit_Linear_RUD_AL}
	\end{subfigure}
	\hfill	
%	\begin{subfigure}[b]{0.49\textwidth}
%		\centering
%		\includegraphics[width=\linewidth]{../images/Function_Fitting/linear_circuitdepth_vs_R2}
%		\caption{}
%		\label{fig:linear_circuitdepth_vs_R2}
%	\end{subfigure}
%	\hfill		
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/sixteenqubit/16qubit_Quadratic_RUD_AL}
		\caption{}
		\label{fig:16qubit_Quadratic_RUD_AL}
	\end{subfigure}
	\hfill
%	\begin{subfigure}[b]{0.49\textwidth}
%		\centering
%		\includegraphics[width=\linewidth]{../images/Function_Fitting/quadratic_circuitdepth_vs_R2}
%		\caption{}
%		\label{fig:quadratic_circuitdepth_vs_R2}
%	\end{subfigure}
%	\hfill		
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/sixteenqubit/16qubit_Sine_RUD_AL.png}
		\caption{}
		\label{fig:16qubit_Sine_RUD_AL}
	\end{subfigure}	
	\hfill
%	\begin{subfigure}[b]{0.49\textwidth}
%		\centering
%		\includegraphics[width=\linewidth]{../images/Function_Fitting/sine_circuitdepth_vs_R2}
%		\caption{}
%		\label{fig:sine_circuitdepth_vs_R2}
%	\end{subfigure}
%	\hfill		
	\caption{(a) linear, (b) quadratic, (c) sine}
	\label{fig:16qubit_RUD_AL}	
\end{figure}

One proposed advantage of using PQCs for ML tasks is that, in some cases, they can outperform classical ML models.\cite{hatakeyama-sato_quantum_2023}
To explore this, we use learning curves with the best performing circuits previously described, where we hold the number of test points to 10\% of the total data points and vary the percentage of the training data.

We compare the best PQCs with the classical models mentioned in subsection \ref{subsection:datasets}
For the linear data using IQP\_Full-Pauli-CRZ  with a RUD and AL of 1, we found that the quantum circuit offers similiar performance

Performance:
Linear
Offers comparable performance to the classical models

quadratic
Outperforms classical linear models (elastic lasso ridge)

sine
Outperforms classical linear models (elastic lasso ridge)

'elastic', 'gpr', 'grad', 'knn', 'krr', 'lasso', 'rfr', 'ridge', 'svr'

@@

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/linear_error_mitigation.png}
		\caption{}
		\label{fig:linear_error_mitigation}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/quadratic_error_mitigation.png}
		\caption{}
		\label{fig:quadratic_error_mitigation}
	\end{subfigure}	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/sine_error_mitigation.png}
		\caption{}
		\label{fig:sine_error_mitigation}
	\end{subfigure}
	\hfill		
	\caption{(a) linear, (b) quadratic, (c) sine function}
	\label{fig:functionfitting_errormitigation}	
\end{figure}

\subsection{BSE}
\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/BSE_heatplots}
		\caption{}
		\label{fig:5BSE_heatplots}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/BSE_boxplots}
		\caption{}
		\label{fig:5BSE_boxplots}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/sixteenqubit/BSE_heatplots}
		\caption{}
		\label{fig:16BSE_heatplots}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/sixteenqubit/BSE_boxplots}
		\caption{}
		\label{fig:16BSE_boxplots}
	\end{subfigure}
	\caption{}
	\label{fig:BSEboxandheat}	
\end{figure}

@@

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/DDCC_heatplots}
		\caption{}
		\label{fig:ddccheatplots}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/DDCC_boxplots}
		\caption{}
		\label{fig:ddccboxplots}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/AllDDCC5_circuitdepth_R2_vs_iterationtime}
		\caption{}
		\label{fig:allddcc5circuitdepthr2vsiterationtime}
	\end{subfigure}	
	\caption{}
	\label{fig:ddcc_all_analysis}	
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{../images/DDCC/DDCC_RUDAL}
	\caption{}
	\label{fig:ddccrudal}
\end{figure}

>>>>

Function fitting 5: all ran with 1000 iterations
Function fitting 16: all ran with 1000 iterations
BSE 5: all ran with 1000 iterations
BSE 16: all ran with 1000 iterations
DDCC

@@

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/sixteenqubit/16qubit_Linear_RUD_AL}
		\caption{}
		\label{fig:16qubit_Linear_RUD_AL}
	\end{subfigure}
	\hfill	
%	\begin{subfigure}[b]{0.49\textwidth}
%		\centering
%		\includegraphics[width=\linewidth]{../images/Function_Fitting/linear_circuitdepth_vs_R2}
%		\caption{}
%		\label{fig:linear_circuitdepth_vs_R2}
%	\end{subfigure}
%	\hfill		
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/sixteenqubit/16qubit_Quadratic_RUD_AL}
		\caption{}
		\label{fig:16qubit_Quadratic_RUD_AL}
	\end{subfigure}
	\hfill
%	\begin{subfigure}[b]{0.49\textwidth}
%		\centering
%		\includegraphics[width=\linewidth]{../images/Function_Fitting/quadratic_circuitdepth_vs_R2}
%		\caption{}
%		\label{fig:quadratic_circuitdepth_vs_R2}
%	\end{subfigure}
%	\hfill		
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/sixteenqubit/16qubit_Sine_RUD_AL.png}
		\caption{}
		\label{fig:16qubit_Sine_RUD_AL}
	\end{subfigure}	
	\hfill
%	\begin{subfigure}[b]{0.49\textwidth}
%		\centering
%		\includegraphics[width=\linewidth]{../images/Function_Fitting/sine_circuitdepth_vs_R2}
%		\caption{}
%		\label{fig:sine_circuitdepth_vs_R2}
%	\end{subfigure}
%	\hfill		
	\caption{The best five (top row), median five (middle row), and worst five (bottom row) circuits for the 16 qubit (a) linear, (b) quadratic, and (c) sine data show that, in general, an increase in the re-upload depth (RUD) or ansatz layers (AL) increases model performance (R$^{2}$ shown on the y-axis).}
	\label{fig:16qubit_RUD_AL}	
\end{figure}

One proposed advantage of using PQCs for ML tasks is that, in some cases, they can outperform classical ML models using less training data.\cite{hatakeyama-sato_quantum_2023}
We examine this phenomena with learning curves using the best performing circuits previously described.
The learning curves are generated by varying the percentage of training data from 10-80\% of the available data, while holding the number of test points to 10\% of the total data.
The three PQCs for each data set are compared to the classical models mentioned in subsection \ref{subsection:datasets} in a statistical manner (left sub-figures in Figs. \ref{fig:linear_learning_curves}, \ref{fig:quadratic_learning_curves}, and \ref{fig:sine_learning_curves}).
Across the full learning curve, IQP\_Full-Pauli-CRZ (linear function fitting) outperforms seven out of nine classical models for the training set and all of the classical models on the test set.
On the quadratic dataset, using IQP\_Full-Pauli-CRX, offers similar performance to the classical models, despite being outperformed on the training set across the learning curve. 
Similar to the IQP\_Full-Pauli-CRZ using the linear data, IQP\_Full-Pauli-CRX outperforms seven out of the nine classical models on the test set across the full learning curve.
On the sine data, A2\_HWE-CZ is consistent across the learning curve, outperforming six of the nine classical models on the training set and all of the models on the test set.
For all three datasets, this implies that the PQCs offer similar performance to classical models that do well on these datasets but provide better results on the test set.

@@

Lastly, using the function fitting datasets we analyze the effects of error mitigation on the model accuracy by comparing state vector simulations with noisy and error mitigated simulations using the \textit{fake\_quebec} backend.
Initially, we explored two different methods of error mitigation zero-noise extrapolation (ZNE) using MITIQ\cite{larose_mitiq_2022} and TREX, as implemented in Qiskit. 
We found that both linear and Richardson ZNE were too costly to simulate on classical computers using 16 qubits for all three function fitting datasets and chose only to use TREX for error mitigation.

IQP\_Full-Pauli-CRZ (linear function fitting)
state vector: train 0.99/test 0.99
TREX: train 0.98/test 0.98
without: train 0.98/test 0.98
\ref{fig:linear_error_mitigation}

IQP\_Full-Pauli-CRX (quadratic function fitting)
state vector: train 0.80/test 0.68
TREX: train 0.69/test 0.60
without: train 0.65/test 0.45
\ref{fig:quadratic_error_mitigation}

A2\_HWE-CZ (sine function fitting)
state vector: train 0.92/test 0.96
TREX: train 0.87/test 0.89
without: train 0.87/test 0.91
\ref{fig:sine_error_mitigation}

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/linear_error_mitigation.png}
		\caption{}
		\label{fig:linear_error_mitigation}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/quadratic_error_mitigation.png}
		\caption{}
		\label{fig:quadratic_error_mitigation}
	\end{subfigure}	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/Function_Fitting/sine_error_mitigation.png}
		\caption{}
		\label{fig:sine_error_mitigation}
	\end{subfigure}
	\hfill		
	\caption{Error mitigation benchmarking for the (a) linear, (b) quadratic, and (c) sine function comparing unmitigated and mitigated error using TREX with state vector simulations.}
	\label{fig:functionfitting_errormitigation}	
\end{figure}

\subsection{BSE}

five qubit \ref{fig:5BSE_heatplots} \ref{fig:5BSE_boxplots}
best encoder-ansatz pair: train R$^{2}$/test R$^{2}$
Best encoder on average train R$^{2}$/test R$^{2}$
Best ansatz on average train R$^{2}$/test R$^{2}$

sixteen qubit \ref{fig:16BSE_heatplots} \ref{fig:16BSE_boxplots}
best encoder-ansatz pair: train R$^{2}$/test R$^{2}$
Best encoder on average train R$^{2}$/test R$^{2}$
Best ansatz on average train R$^{2}$/test R$^{2}$

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/BSE_heatplots}
		\caption{}
		\label{fig:5BSE_heatplots}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/fivequbit/BSE_boxplots}
		\caption{}
		\label{fig:5BSE_boxplots}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/sixteenqubit/BSE_heatplots}
		\caption{}
		\label{fig:16BSE_heatplots}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/BSE/sixteenqubit/BSE_boxplots}
		\caption{}
		\label{fig:16BSE_boxplots}
	\end{subfigure}
	\caption{}
	\label{fig:BSEboxandheat}	
\end{figure}

\ref{fig:bse5rudal}

\ref{fig:bse16rudal}

@@

A2\_HWE-CNOT Train R$^{2}$ 0.62/test R$^{2}$ 0.62 
Best encoder average Train R$^{2}$ X/test R$^{2}$ Y
Best ansatz HWE-CNOT average Train R$^{2}$ X/test R$^{2}$ Y

\begin{figure}[H]
	\centering	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/DDCC_heatplots}
		\caption{}
		\label{fig:ddccheatplots}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.49\textwidth}
		\centering
		\includegraphics[width=\linewidth]{../images/DDCC/DDCC_boxplots}
		\caption{}
		\label{fig:ddccboxplots}
	\end{subfigure}
	\hfill
%	\begin{subfigure}[b]{0.49\textwidth}
%		\centering
%		\includegraphics[width=\linewidth]{../images/DDCC/AllDDCC5_circuitdepth_R2_vs_iterationtime}
%		\caption{}
%		\label{fig:allddcc5circuitdepthr2vsiterationtime}
%	\end{subfigure}	
	\caption{}
	\label{fig:ddcc_all_analysis}	
\end{figure}

(AL,RUD)=(1,1) Train R$^{2}$ 0.62/test R$^{2}$ 0.62
(AL,RUD)=(1,3) Train R$^{2}$ 0.85/test R$^{2}$ 0.85
(AL,RUD)=(1,5) Train R$^{2}$ 0.82/test R$^{2}$ 0.83
(AL,RUD)=(3,1) Train R$^{2}$ 0.71/test R$^{2}$ 0.71
(AL,RUD)=(5,1) Train R$^{2}$ 0.77/test R$^{2}$ 0.77

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{../images/DDCC/DDCC_RUDAL}
	\caption{Model evaluation, using R$^{2}$ (y-axis), of re-upload depths (RUD) and ansatz layers (AL) of 1, 3, and 5 for the A2\_HWE-CNOT using the DDCC dataset. The left side of the plot denotes the training set and the right side the test set.}
	\label{fig:ddccrudal}
\end{figure}
